Prototype city generation for mobility simulation in urban typologies: a generalized framework

### 2017-12-27 manuscript/manuscript.tex
reorganized document

\section{Introduction}
Household assignment and spatial allocation are two big issues we need to solve when facing the problem of population synthesis.
Specifically, the first one addresses the problem of how to match individuals to the household where they belong, while the spatial allocation distributes households in geographical space.
It is necessary because the ultimate goal of our work is to do transportation simulation, which can only be achieved by knowing the location of each agent. \\

\section{Background}
There has been lots of literature on population synthesis.
The methods used varies from Iterative Proportional Fitting (IPF), MCMC-based sampling \cite{farooq2013simulation}, to Bayesian network approach \cite{sun2015bayesian}. The original work in this field is presented in \cite{beckman1996creating}. It used a method called Iterative Proportional Fitting (IPF) to synthesize population based on data from two sources. What they do is iteratively fitting the cells of a multi-way contingency table expanded by attributes that characterize a household/individual so that it satisfies the constraints of marginal distribution of demographics and also preserves the correlation structure carried by sample data of the whole population. Here the marginals of demographics comes from census data in the level of census tract or block group while the correlation structure need to be derived from survey data of target population. In case of the absence of disaggregated sample data, IPF can still give a solution by iteratively adjust the cells in accordance only with the marginal constraints. However, lots of different joint distributions can share the same marginals and IPF may converge to anyone of them. This is certainly a downside of traditional IPF, and is also the reason why we need the correlation structure of sample data to make the solution captures the true population distribution. After the fitting of the table, a desired number of synthetic population is sampled according to individual or household weights in the table. And this is where the sensitivity of IPF to the quality of data and the sample size stems from. In this method, spatial allocation is solved concurrently with the population synthesis, since it generates the population for each zone and then adds up to the whole population. On contrary to this, one can also synthesize the total population first and then allocate them to different zones. However, this includes implementing a detailed matching method between households and zones, like what's developed in \cite{ge2014virtual}, which is also very time-consuming.

@@

\section{Data}
Most paper in population synthesis used Public Use Microdata Sample (PUMS) data, which is available both in individual level and household level.
It includes multiple socio-economic parameters of a household or an individual in the region.
The sample rate of this dataset is about 1\% to 10\% of the total population in the region. 
PUMS data is very complete in US cities and can be easily downloaded from government census website \footnote{https://www.census.gov/programs-surveys/acs/data/pums.html}, but the same data for cities in developing countries can be relatively outdated and harder to find.
The sample data is used as seeds to initialize the contingency table in IPF methods.
Another important part of data is called aggregate data \footnote{https://factfinder.census.gov}.
They give marginal distribution for variables of our interests.
It's worth noticing that the marginal distributions for subsets with more than 2 variables can be rarely found.
An ideal format for aggregate data is that we have the marginal distributions in the level of geographical zone cause IPF is carried out separately in each zone.
So when downloading the aggregate data, one should first choose a geographic type in corresponding to the zones in the virtual city, then specify the variables or tables you want.

@@

\subsection{Network processing}

\section{Method}
\begin{table}[h!]
  \centering
  \begin{tabular}{l l l}
    &&
  \end{tabular}
  \caption{List of symbols for variables and parameters}
  \label{tab:symbols}
\end{table}

\begin{table}[h!]
  \centering
  \begin{tabular}{l l l}\toprule
    \bf SN &\bf Name &\bf Supply candidate \\\midrule
    1 & Auto Sprawl & Baltimore \\
    2 & Innovative Heavyweights & Madrid \\
    3 & Sustainable Anchors & Vancouver \\
    4 & Congested Middleweights & Bandung \\
    5 & Boomers & Lagos \\
    6 & Emerging Green Anchors & Hiroshima \\
    7 & Dense Clique & Isfahan\\
    8 & Hybrid Congested & Mashdad \\
    9 & BRT-{\it STAR}s & Sao Paulo  \\
    10 & Metro-Bike Giants & Dalian \\ \bottomrule
  \end{tabular}
  \caption{Prototype city descriptions and candidates}
  \label{tab:proto}
\end{table}

\begin{table}[h!]
  \centering
  \begin{tabular}{l l l l l }
    City/Typology & Microdata & Aggregate data & Land use & Transit \\
  \end{tabular}
  \caption{Data availability for candidate cities}
  \label{tab:avail}
\end{table}

\subsection{Supply}

@@

    \node[] (utp-location) [right= of utp-source] {};
    \node[] (sal-location) [right= of sal-source] {};
    \node[loca] (taz-location) [right= of taz-source] {Work (WORK), Education (EDU)};
    \node[loca] (cell-location)[right= of cell-source]{Household (HH)};
  \end{tikzpicture}
  \caption{The four tiers of the prototype city development framework. The UTP represents the urban typology, and is given by the metro area of the candidate city chosen for the supply.
  The Second Administrative Levels are the highest-level regions into which the UTP is subdivided. TAZ's are obtained for the UTP given the candidate road network. Where unavailable, these are generated as Thiessen polygons. CELLs are obtained by gridding over the SALs (or TAZs). The corresponding land use weight at the CELL level is used to determine the  {\it number} and {\it location} of the households therein; as well as the {\it numbers} of WORK and EDU points in each TAZ.}
  \label{fig:framework}
\end{figure}
\subsection{Geolocation and spatial assignment}
\textcite{moekel2003microsimulation} present an intuitive approach to assigning households to zones, which we propose to implement.
First, they obtain land-use data, which they use to disaggregate zonal characteristics (e.g.\ household location or firm location) into raster cells.
The size of these cells are chosen based on computational tractability and desired resolution.
Weights are then assigned to the cells based on the corresponding land-use category.
For the household example, \textcite{moekel2003microsimulation} aggregate land-use attributes into the following: high-density housing, low-density housing, industry and open space;
to which weights of 10, 5, 1 and 1 are assigned, respectively.
The weights are then normalized by zone, and households are randomly selected to populate each cell to fit each proportion.
The address of each household is assumed to be the centroid of each cell, which each associated with its parent zone.
The rasterization is performed by a so-called ``point-in-polygon'' algorithm.
This approach presents a fast way for household assignment, assuming prior zonal assignment is available.
Our implementation is outlined in the following subsections.

>>>>

\section{Introduction}
We describe in detail an automated pipeline for generating urban typology prototypes.
These prototypes are simulation-ready environments that represent outcomes for the urban typologies discovered from global data collected in the first phase of the Mobility of the Future project.
The organization of this document is as follows: First, a brief background and motivation is provided in Section 2. Subsequently, we describe our data sources and organization.
In Section 4, the overall framework and submethods are described, with relationships clearly defined.
Results are then presented and discussed in Section 5.
Finally, we conclude with a summary and layout steps for further work.

The contribution of this work lies in its highly generalized approach and its reliance only on open data.
There are several examples in the extant literation on the synthesis and generation of cities for agent-based mobility simulation.
However, these are mostly city-specific.

Here, we combine existing methods in creating a pipeline that produces prototypes that represent the heterogeneity of urban structure and outcomes.
The outputs from our approach serve as inputs in our state-of-the-art integrated urban laboratory---SimMobility \citep{adnan2016simmobility}, which we will use to evaluate future scenarios across these prototypes.

\section{Background}
{\it Too population-synthesis-heavy!! Needs REVISION; some material might need to go to the appendix}\\

There has been lots of literature on population synthesis.
The methods used varies from Iterative Proportional Fitting (IPF), MCMC-based sampling \cite{farooq2013simulation}, to Bayesian network approach \cite{sun2015bayesian}. The original work in this field is presented in \cite{beckman1996creating}. It used a method called Iterative Proportional Fitting (IPF) to synthesize population based on data from two sources. What they do is iteratively fitting the cells of a multi-way contingency table expanded by attributes that characterize a household/individual so that it satisfies the constraints of marginal distribution of demographics and also preserves the correlation structure carried by sample data of the whole population. Here the marginals of demographics comes from census data in the level of census tract or block group while the correlation structure need to be derived from survey data of target population. In case of the absence of disaggregated sample data, IPF can still give a solution by iteratively adjust the cells in accordance only with the marginal constraints. However, lots of different joint distributions can share the same marginals and IPF may converge to anyone of them. This is certainly a downside of traditional IPF, and is also the reason why we need the correlation structure of sample data to make the solution captures the true population distribution. After the fitting of the table, a desired number of synthetic population is sampled according to individual or household weights in the table. And this is where the sensitivity of IPF to the quality of data and the sample size stems from. In this method, spatial allocation is solved concurrently with the population synthesis, since it generates the population for each zone and then adds up to the whole population. On contrary to this, one can also synthesize the total population first and then allocate them to different zones. However, this includes implementing a detailed matching method between households and zones, like what's developed in \cite{ge2014virtual}, which is also very time-consuming.

@@

\textcite{moekel2003microsimulation} present an intuitive approach to assigning households to zones, which we propose to implement.
First, they obtain land-use data, which they use to disaggregate zonal characteristics (e.g.\ household location or firm location) into raster cells.
The size of these cells are chosen based on computational tractability and desired resolution.
Weights are then assigned to the cells based on the corresponding land-use category.
For the household example, \textcite{moekel2003microsimulation} aggregate land-use attributes into the following: high-density housing, low-density housing, industry and open space;
to which weights of 10, 5, 1 and 1 are assigned, respectively.
The weights are then normalized by zone, and households are randomly selected to populate each cell to fit each proportion.
The address of each household is assumed to be the centroid of each cell, which each associated with its parent zone.
The rasterization is performed by a so-called ``point-in-polygon'' algorithm.
This approach presents a fast way for household assignment, assuming prior zonal assignment is available.

\section{Data}
{\it NEEDS REVISION!}\\
Most paper in population synthesis used Public Use Microdata Sample (PUMS) data, which is available both in individual level and household level.
It includes multiple socio-economic parameters of a household or an individual in the region.
The sample rate of this dataset is about 1\% to 10\% of the total population in the region. 
PUMS data is very complete in US cities and can be easily downloaded from government census website \footnote{https://www.census.gov/programs-surveys/acs/data/pums.html}, but the same data for cities in developing countries can be relatively outdated and harder to find.
The sample data is used as seeds to initialize the contingency table in IPF methods.
Another important part of data is called aggregate data \footnote{https://factfinder.census.gov}.
They give marginal distribution for variables of our interests.
It's worth noticing that the marginal distributions for subsets with more than 2 variables can be rarely found.
An ideal format for aggregate data is that we have the marginal distributions in the level of geographical zone cause IPF is carried out separately in each zone.
So when downloading the aggregate data, one should first choose a geographic type in corresponding to the zones in the virtual city, then specify the variables or tables you want.

@@

\subsection{Urban typologies}
\begin{table}[h!]
  \centering
  \caption{Prototype city descriptions and candidates}
  \label{tab:proto}
  \begin{tabular}{l l l}\toprule
    \bf SN &\bf Name &\bf Supply candidate \\\midrule
    1 & Auto Sprawl & Baltimore \\
    2 & Innovative Heavyweights & Madrid \\
    3 & Sustainable Anchors & Vancouver \\
    4 & Congested Middleweights & Bandung \\
    5 & Boomers & Lagos \\
    6 & Emerging Green Anchors & Hiroshima \\
    7 & Dense Clique & Isfahan\\
    8 & Hybrid Congested & Mashdad \\
    9 & BRT-{\it STAR}s & Sao Paulo  \\
    10 & Metro-Bike Giants & Dalian \\ \bottomrule
  \end{tabular}
\end{table}

\begin{table}[h!]
  \centering
  \caption{Data availability for candidate cities}
  \label{tab:avail}
  \begin{tabular}{l l l l l }
    City/Typology & Microdata & Aggregate data & Land use & Transit \\
  \end{tabular}
\end{table}

\section{Method}

@@

    \node[] (utp-location) [right= of utp-source] {};
    \node[] (sal-location) [right= of sal-source] {};
    \node[loca] (taz-location) [right= of taz-source] {Work (WORK), Education (EDU)};
    \node[loca] (cell-location)[right= of cell-source]{Household (HH)};
  \end{tikzpicture}
  \caption{The four tiers of the prototype city development framework. The UTP represents the urban typology, and is given by the metro area of the candidate city chosen for the supply.
  The Second Administrative Levels are the highest-level regions into which the UTP is subdivided. TAZ's are obtained for the UTP given the candidate road network. Where unavailable, these are generated as Thiessen polygons. CELLs are obtained by gridding over the SALs (or TAZs). The corresponding land use weight at the CELL level is used to determine the  {\it number} and {\it location} of the households therein; as well as the {\it numbers} of WORK and EDU points in each TAZ.}
  \label{fig:framework}
\end{figure}

\begin{table}[h!]
  \centering
  \begin{tabular}{l l l}
    &&
  \end{tabular}
  \caption{List of symbols for variables and parameters}
  \label{tab:symbols}
\end{table}

\subsection{Supply}
The supply pipeline can be summarized in the following steps:
\begin{itemize}
\item Obtain candidate city for urban typology based on that closest to centroid of supply factors
\item Process road network from supply candidate and generate SimMobility inputs
\item Generate public transit graph from GTFS and other sources
\item Process land-use data polygons
\end{itemize}

\subsection{Geolocation and spatial assignment}
Our implementation is outlined in the following subsections.
